---
title: "EA Challenge"
author: "Jose Manuel Vera"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: Cayman
    highlight: github
---

List of required libraries. System info at the bottom of this file.


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)

```

```{r libraries}

rm(list=ls())

library(rio)
library(broom)
library(knitr)
library(plotly)
library(RColorBrewer)
library(kableExtra)
library(summarytools)
library(prettydoc)
library(jsonlite)
library(maps)
library(htmltools)
library(hrbrthemes)
library(DataExplorer)
library(tidyverse)
library(viridis)
```

## DATA LOAD

Importing all CSV files with just one command and clean some quirks

```{r csvload}
mydata <-
    rio::import_list(dir("raw_data/",full.names = TRUE,pattern = "*.csv"),
                     rbind = FALSE)

# removing redundant data: "CompleteDataset"
mydata <- mydata[names(mydata) != "CompleteDataset"] 

# Force numeric all numeric columns in Player Attribute Data
mydata$PlayerAttributeData[]  <- lapply(mydata$PlayerAttributeData, function(x) as.numeric(as.character(x)))

# Remove index/rowname columns in all dataframes
mydata <- lapply(mydata, function(x) { x["V1"] <- NULL; x })

# Personal Data has an "Unnamed:0" Column. Get rid of it (First Col)
mydata$PlayerPersonalData <- mydata$PlayerPersonalData[2:ncol(mydata$PlayerPersonalData)]


# Extracting dataframes from list to global environment
# (This is optional. Working with list of objects is easier in R.
# Executed here just for easier workflow comprehension)
list2env(mydata, globalenv()) 

# List of data frame objects
ls()[sapply(ls(), function(x) class(get(x))) == 'data.frame']
```

# OBJECTIVE 1 

## Loading and merging Country & Language Data

```{r fromJSON, echo=TRUE, warning=FALSE}

countries <- fromJSON("raw_data/countries.json") %>%
            data.table::rbindlist() %>%
            as_tibble() %>%
            mutate(languages = map(languages, ~ replace(.x, is.null(.x), "unknown"))) %>%  # avoid empty list items
            mutate(languages = unlist(languages)) %>%
            rename(lang_code = languages, country_name = name, native_name = native)



languages <- tibble(col = fromJSON("raw_data/languages.json")) %>%
             unnest_wider(col) %>%
             mutate(languages = names(fromJSON("raw_data/languages.json"))) %>%
             rename(lang_code = languages, lang_name = name, native_lang = native)



continents <- fromJSON("raw_data/continents.json") %>%
              unlist() %>%
              stack() %>%
              rename(continent = ind, continent_name = values)


final_languages_csv <-
              inner_join(countries, languages, by = "lang_code") %>%
              inner_join(continents, by = "continent") %>%
              select(-rtl)


# WRITING CSV TO DISK ----
rio::export(final_languages_csv,"processed_data/final_languages.csv")


# Look at the csv
kable(final_languages_csv) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    font_size = 12
  ) %>%
  row_spec(row = 0, color = "#660033") %>% 
  scroll_box(width = "100%", height = "400px")

```

## Saving to database (Optional)

When data is not huge I usually tend to work with MySQL. A database I know very well because I worked many years in the past with different versions of it.
MySQL package from R is deprecated and MariaDB should be used instead.

**Uncomment and set credentials to make it work (be sure the user has Write privileges)**

```{r database, echo=TRUE}

# mycon <-
#   dbConnect(RMariaDB::MariaDB(),
#             user = 'user',
#             password = 'pass',
#             dbname = 'dbname',
#             host = 'database.host'
#   )
# 
# on.exit(dbDisconnect(mycon), add = TRUE)
# 
# dbExecute(mycon, as.character('set character set "utf8"'))
# 
# copy_to(destination, final_languages_csv, name = "languages_table", overwrite = TRUE)

```

## Basic DataSet Description

```{r introduce}

 lapply(mydata, DataExplorer::introduce) %>% 
                dplyr::bind_rows() %>% 
                dplyr::bind_cols(tibble(names(mydata)) %>% rename("DataSet"=1)) %>%
                dplyr::select(DataSet, everything()) %>% 
                kable(caption = "Basic Data Description") %>% 
                kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10) %>% 
               row_spec(row = 0, color = "#660033") %>% 
               scroll_box(width = "100%", height = "300px")


```

## Basic Summaries

```{r summarytable, echo=TRUE, results='asis'}

# function to return a summary table
get_summary_table <- function(df) {
  
  df %>% 
    descr() %>% 
    kable(caption = "Summary Table") %>% 
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10) %>% 
    row_spec(row = 0, color = "#660033") %>% 
    scroll_box(width = "100%", height = "300px")
}



lapply(mydata, function(x) get_summary_table(x)) 

```

## Checking for empty values (Percentage)

```{r NAsPercent}
lapply(mydata, function(x) naniar::gg_miss_var(x,show_pct = TRUE)) 
```

## Seeking if any Column contains dates

**Spoiler: No dates found in any column.**

```{r seekingDates}

date_seek <- lapply(mydata, function(x) {
  x[] <- lapply(x, lubridate::is.Date)
  x
})

 sapply(date_seek, function(x) any(unlist(x)))

```

# Distributions: Player Attributes

```{r PlayerAttributesDist}
plot_histogram(mydata$PlayerAttributeData)
```


# Distributions: Player Personal Data

```{r PlayerPersonalDist}
plot_histogram(mydata$PlayerPersonalData)
```

# Distributions: Player Position Data

```{r PlayerPositionDist}
plot_histogram(mydata$PlayerPlayingPositionData)
```

## EDA: Players & Languages

A table summarizing player and languages is supposed they speak.

```{r edaJoin}

players_languages <-
  inner_join(final_languages_csv,
             PlayerPersonalData,
             by = c("country_name" = "Nationality"))  %>%
  select(lang_name, Club, country_name, continent_name, Name) %>%
  group_by(Name) %>%
  summarise(
    languages = toString(unique(lang_name)),
    club =  first(Club),
    country = first(country_name),
    continent = first(continent_name)
  )

kable(players_languages) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    font_size = 12
  ) %>%
  row_spec(row = 0, color = "#660033") %>% 
  scroll_box(width = "100%", height = "400px")

```

## EDA: Top N in polyglot countries

```{r fig.width=8}
top_languages <- players_languages %>%
  filter(grepl(",", languages, fixed = TRUE)) %>% 
  group_by(languages,country) %>% 
  tally() %>%
  arrange(desc(n)) %>% 
  head(15)


  ggplot(top_languages,aes(x = reorder(languages, n), y = n, fill=country, palette="dark2")) +
   geom_bar(stat = "identity")+
   geom_text(aes(label=n), vjust=0.5, hjust=1, color="white", size=4)+
   xlab("") +
   coord_flip() +
   theme(plot.title = element_text(color = "black", size = 24, face = "bold"),
         plot.subtitle = element_text(size=18))+
   labs(title='Top 15 languages in polyglot countries') +
    labs(subtitle="Color by Country")

```

## EDA: Top N global 

```{r fig.width=6}
top_languages <- inner_join(final_languages_csv,
             PlayerPersonalData,
             by = c("country_name" = "Nationality"))  %>%
  select(lang_name, Club, country_name, continent_name, Name) %>%
  group_by(lang_name) %>% 
  tally() %>%
  arrange(desc(n)) %>% 
  head(15)


  ggplot(top_languages,aes(x = reorder(lang_name, n), y = n)) +
   geom_bar(stat = "identity",fill="darkviolet" )+
   geom_text(aes(label=n), vjust=0.5, hjust=1, color="white", size=4)+
   xlab("") +
   coord_flip() + 
    theme( plot.title = element_text(color = "black", size = 24, face = "bold"))+
   labs(title='Top 15 languages all around the world')
  
```


# OBJECTIVE 2

Build a Data Visualization of the data FIFA game by country and languages. Clustering to help us for take the best
decisions for Localization.
Tasks:
• Faceting, adding visual variables, and summarization
• Profile the different player and teams.

## Players by Country

```{r continents, fig.width=12,fig.height=12}

# Loading geographical data per player

players_by_country <- 
             PlayerPersonalData %>%
  group_by(Nationality) %>%
  tally() %>%
  magrittr::set_colnames(c("region","value"))%>% 
  ungroup() %>% 
   mutate(
    region = ifelse(region == "United States", "USA", region),
    region = ifelse(region == "DR Congo", "Democratic Republic of the Congo", region),
    region = ifelse(region == "China PR", "China", region),
    region = ifelse(region == "Congo", "Democratic Republic of the Congo", region),
    region = ifelse(region == "Bosnia Herzegovina", "Bosnia and Herzegovina", region),
    region = ifelse(region == "Scotland", "UK", region),
    region = ifelse(region == "England", "UK", region),
    region = ifelse(region == "Northern Ireland", "Ireland", region),
    region = ifelse(region == "Republic of Ireland", "Ireland", region),
    region = ifelse(region == "Central African Rep.", "Central African Republic", region),
    region = ifelse(region == "Wales", "UK", region)
    )  


# loading world maps data

world_map <- map_data("world")

# joining map data with players data

players_map <- left_join(players_by_country, world_map, by = "region") %>% 
  rename(players=value)

players_map$hover <- with(players_map, paste0(region,"<br>",players))

g <- ggplot(players_map, aes(long, lat, group = group, text=hover))+
  geom_polygon(aes(fill = players), color = "white")+
   scale_fill_viridis_c(option = "C") +
   labs(title='# Players by Continent / Country')+
  hrbrthemes::theme_ipsum() 


ggplotly(g,tooltip="text")


```

# Top languages per Continent: Faceted version

```{r continentamount, fig.width=10}

continent_facets <- players_languages %>%
  group_by(continent,languages) %>% 
  tally() %>%
  arrange(desc(n)) %>% 
  top_n(n = 5) 


ggplot(continent_facets,
       aes(
         x = reorder(languages, n),
         y = n,
         fill = languages,
         palette = "dark2"
       )) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = n),
    vjust = 0.5,
    hjust = 1,
    color = "black",
    size = 4
  ) +
  xlab("") +
  coord_flip() +
  theme_minimal()+
  theme(
    legend.position = "none",
    plot.title = element_text(color = "black", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10)
  ) +
  facet_grid(~ continent, scales = "free", space = "free")
  
  
```


# Languages per continent (non faceted version)

```{r asialang, fig.width=6}
continent_facets %>% filter(continent=="Asia") %>% 
  ggplot( aes(
         x = reorder(languages, n),
         y = n,
         fill = languages,
         palette = "dark2"
       )) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = n),
    vjust = 0.5,
    hjust = 1,
    color = "white",
    size = 4
  ) +
  xlab("") +
  coord_flip() +
  theme_minimal()+
  theme(
    legend.position = "none",
    plot.title = element_text(color = "black", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10)
  ) +
  labs(title = "TOP 5 Asiatic Players Languages")

```
```{r EUROlang, fig.width=5}
continent_facets %>% filter(continent=="Europe") %>% 
  ggplot( aes(
         x = reorder(languages, n),
         y = n,
         fill = languages,
         palette = "dark2"
       )) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = n),
    vjust = 0.5,
    hjust = 1,
    color = "white",
    size = 4
  ) +
  xlab("") +
  coord_flip() +
  theme_minimal()+
  theme(
    legend.position = "none",
    plot.title = element_text(color = "black", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10)
  ) +
  labs(title = "TOP 5 Europe Players Languages")

```
```{r Africalang, fig.width=6}
continent_facets %>% filter(continent=="Africa") %>% 
  ggplot( aes(
         x = reorder(languages, n),
         y = n,
         fill = languages,
         palette = "dark2"
       )) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = n),
    vjust = 0.5,
    hjust = 1,
    color = "white",
    size = 4
  ) +
  xlab("") +
  coord_flip() +
  theme_minimal()+
  theme(
    legend.position = "none",
    plot.title = element_text(color = "black", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10)
  ) +
  labs(title = "TOP 5 African players languages")

```

```{r southamlang, fig.width=6}
continent_facets %>% filter(continent=="South America") %>% 
  ggplot( aes(
         x = reorder(languages, n),
         y = n,
         fill = languages,
         palette = "dark2"
       )) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = n),
    vjust = 0.5,
    hjust = 1,
    color = "white",
    size = 4
  ) +
  xlab("") +
  coord_flip() +
  theme_minimal()+
  theme(
    legend.position = "none",
    plot.title = element_text(color = "black", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10)
  ) +
  labs(title = "TOP 5 South American players languages")

```


```{r northamlang, fig.width=6}
continent_facets %>% filter(continent=="North America") %>% 
  ggplot( aes(
         x = reorder(languages, n),
         y = n,
         fill = languages,
         palette = "dark2"
       )) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = n),
    vjust = 0.5,
    hjust = 1,
    color = "white",
    size = 4
  ) +
  xlab("") +
  coord_flip() +
  theme_minimal()+
  theme(
    legend.position = "none",
    plot.title = element_text(color = "black", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10)
  ) +
  labs(title = "TOP 5 North American players languages")

```


```{r oceanialang, fig.width=6}
continent_facets %>% filter(continent=="Oceania") %>% 
  ggplot( aes(
         x = reorder(languages, n),
         y = n,
         fill = languages,
         palette = "dark2"
       )) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = n),
    vjust = 0.5,
    hjust = 1,
    color = "white",
    size = 4
  ) +
  xlab("") +
  coord_flip() +
  theme_minimal()+
  theme(
    legend.position = "none",
    plot.title = element_text(color = "black", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10)
  ) +
  labs(title = "TOP 5 Oceania players languages")

```